<!-- R Commander Markdown Template -->

Replace with Main Title
=======================

### Your Name

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```


```{r echo=FALSE}
# include this code chunk as-is to enable 3D graphs
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```


```{r}
names(marvel_dc) <- make.names(names(marvel_dc))
```

```{r}
summary(marvel_dc)
```


```{r}
scatterplot(APPEARANCES~page_id | group, regLine=FALSE, smooth=FALSE, 
  boxplots=FALSE, by.groups=TRUE, data=marvel_dc)
```

```{r}
library(rgl, pos=67)
```


```{r}
library(nlme, pos=68)
```


```{r}
library(mgcv, pos=68)
```


```{r, webgl=TRUE}
scatter3d(APPEARANCES~page_id+YEAR|group, data=marvel_dc, surface=FALSE, 
  residuals=TRUE, parallel=FALSE, bg="white", axis.scales=TRUE, grid=TRUE, 
  ellipsoid=FALSE)
```


```{r, webgl=TRUE}
GLM.2 <- glm(group ~ YEAR +GSM + EYE + HAIR, family=binomial(logit), 
  data=marvel_dc)
summary(GLM.2)
exp(coef(GLM.2))  # Exponentiated coefficients ("odds ratios")
```

```{r}
library(lattice, pos=70)
```

```{r}
load("C:/Users/rkrishnan/Documents/01 Personal/MS/SCM 651/week 7/Business+Analytics+-+Week+7+oj.csv")
```


```{r}
data()

```{r}
Orange <- 
  read.table("C:/Users/rkrishnan/Documents/01 Personal/MS/SCM 651/week 7/Business+Analytics+-+Week+7+oj.csv",
   header=TRUE, sep="", na.strings="NA", dec=".", strip.white=TRUE)
```

```{r}
summary(Orange)
```


```{r}
load("C:/Users/rkrishnan/Documents/01 Personal/MS/SCM 651/week 7/Business+Analytics+-+Week+7+oj.csv")
```


```{r}
Dataset <- 
  read.table("C:/Users/rkrishnan/Documents/01 Personal/MS/SCM 651/week 7/Business+Analytics+-+Week+7+oj.csv",
   header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)
```

